Course Project 1
======================================================
### Installing the different packages that may be required in the course project
```{r}
install.packages("ggplot2")
install.packages("plyr")
install.packages("dplyr")

library(ggplot2)
library(plyr)
library(dplyr)
```
### Loading the dataset

```{r echo=TRUE}
counts <- read.csv("C:/Users/amank/Downloads/reproducible_week2/activity.csv", na.strings = "")
summary(counts)
str(counts)
```

##Total number of steps per day
```{r echo=TRUE}
counts$day <- as.Date(counts$date)
counts$DateTime <- as.POSIXct(counts$date,format = "%Y-%m-%d")
clean <- counts[!is.na(counts$steps),]
clean
```
## Converting the steps into numeric

```{r echo=TRUE}
counts$steps <- as.numeric(as.character(counts$steps))
total <- aggregate(counts$steps ~ counts$date, FUN = sum,)
colnames(total) <- c("Date","Steps")
```
## Plotting the graph

```{r echo=TRUE}
hist(total$Steps, breaks=5, xlab="Steps", main = "Total Steps per Day")
```

##Mean and median of the total number of steps

```{r echo=TRUE}

as.integer(mean(total$Steps))
as.integer(median(total$Steps))
```

## Average Daily Pattern Activity
```{r echo=TRUE}
clean <- counts[!is.na(counts$steps),]
average_pattern <- ddply(clean, .(interval), summarize, Avg = mean(steps))
a <- ggplot(data = average_pattern, aes(x=interval, y=Avg), xlab = "Interval", ylab="Average Count of Steps") +
geom_line()+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Average Number of Steps per Interval")

maximum <- max(average_pattern$Avg)
average_pattern[average_pattern$Avg==maximum,1]
```

##Inputing missing values
```{r echo=TRUE}
nrow(counts[is.na(counts$steps),])

counts2<- counts
nas<- is.na(counts2$steps)
average_difference <- tapply(counts2$steps, counts2$interval, mean, na.rm=TRUE, simplify = TRUE)
counts2$steps[nas] <- average_difference[as.character(counts2$interval[nas])]
names(counts2)

counts2 <- counts2[,c("date","interval","steps")]

#####creating the histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

total_steps<- counts2%>%
  group_by(date)%>%
  summarise(number_steps = sum(steps, na.rm= TRUE))
total_steps

ggplot(total_steps, aes(x = number_steps)) +
        geom_histogram(fill = "black", binwidth = 600) +
        labs(title = "Daily Number of Steps including Missing values", x = "Interval", y = "Number of Steps")
##### Calculating the mean and median after imputing the missing values
as.integer(mean(total_steps$number_steps))
as.integer(median(total_steps$number_steps))
```

##Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
counts2<- counts2%>%
        mutate(weektype= ifelse(weekdays(counts2$date)=="Saturday" | weekdays(counts2$date)=="Sunday", "Weekend", "Weekday"))


total_steps_2<- counts2%>%
        group_by(interval, weektype)%>%
        summarise(number_steps_2 = mean(steps, na.rm=TRUE))
total_steps_2

b <- ggplot(total_steps_2, aes(x =interval , y=number_steps_2, color=weektype)) +
       geom_line() +
       labs(title = "Average of daily Steps by Weektype", x = "Interval", y = "Number of Steps") +
       facet_wrap(~weektype, ncol = 1, nrow=2)
print(b)
```